A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2020-06-22 03_Prep_data.Rmd
Move AUDPC plotting to Preliminary analysis
Change headings order to improve flow

The trials `r unique(PM_MB_means[is.na(PM_MB_means$AUDPC_m),"trial_ref"]) `Some trials don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r}
source(here("R/AUDPC_for_single_obs.R"))

@@

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(trial_ref, location, Y_error_type, fungicide_ai) %>%
   summarise(DMI_treatments = length(Y_error_type)) %>%
   arrange(Y_error_type) %>%
   select(!fungicide_ai)%>%
   kable(caption = "Number of treatments for each trial and the error type reported", align = "c") %>%
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")
```

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[1] * (LSD /
                                                                             Tcrit) ^ 2) / 2 # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates))
      } # reset Sv vector if this is the first loop

      Sv[j] = V_yield /
         PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j]

@@

   # Replace the values in the dataset
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$Y_error_type <-
      "SamplingVariance"

@@

@@

Using the model without dose produces a better fit, as shown by the lower AIC and is significantly different from the model `m1`. 
We will remove the dose term from the trial grouping.

```{r}
slimmer_PM_dat$trial <- slimmer_PM_dat$trial_noDose
slimmer_PM_dat <- slimmer_PM_dat[,colnames(slimmer_PM_dat) != "trial_noDose"] #remove second trial column
write.csv(slimmer_PM_dat, file = "cache/slimmer_PM_clusterdat.csv", row.names = FALSE)
```

Our data currently uses sample variance to define the yield error. We should expect that when yield increases so does the variance. Let's quickly plot this to check.

```{r Var_mean_sqaures}
# redefine sample variance calculated from mean square
slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare /
   (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^ 2)

slimmer_PM_dat %>%
   ggplot(aes(x = grain_yield.t.ha, y = vi)) +
   geom_point()
```

## AUDPC visualise raw data

>How well does 'area under the disease progress curve' (AUDPC) predict grain yield?

and _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

We would expect to see grain yield decline with increasing AUDPC. Let's just plot yield vs AUDPC.

```{r Yield_vs_AUDPC}

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))

There appears to be a weak positive relationship between the two continuous variables.  
Does a linear model confirm this?

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

Lets tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship. 
Let's do the plot again for each trial where there are no missing values.  
Note: The plots are ordered by mean yield of the trial.

```{r AUDPCxTrial}
sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_means$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

This sequence of graphs show that AUDPC and disease incidence tends not impact on yields if the season is a low yielding, ie lower than 1 tonne per hectare.  
The higher the yield the greater greater the influence on the disease on the yield especially if the crop is high yielding with AUDPC also being large enough to reduce yield.

Let's have a look again at the same sequence of graphs and analyses for area under the disease progress _stairs_.

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

summary(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i,]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control", ],
          col = "red")

```
The plots from the AUPDS is similar to the AUDPC, however according to the linear model the AUDPC is a better, significant fit.

A mixed effect model would be more suitable to investigate the interactive effects of location, season, and trial. 
The mixed effect model will help us decide on the best approach and structure for our meta-analysis model.

>>>>

@@

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(trial_ref, location,year ,Y_error_type, fungicide_ai) %>%
   summarise(DMI_treatments = length(Y_error_type)) %>%
   arrange(Y_error_type) %>%
   select(!fungicide_ai)%>%
   kable(caption = "Number of treatments for each trial and the error type reported", align = "c") %>%
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")
```

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[1] * (LSD /
                                                                            Tcrit) ^ 2) / 2 # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates))
      } # reset Sv vector if this is the first loop

      Sv[j] = V_yield /
         PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[j]

@@

   # Replace the values in the dataset
   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Y_error_type <-
      "SamplingVariance"

@@

### Check variance of data-set
Our data currently uses sample variance to define the yield error. We should expect that when yield increases so does the variance. Let's quickly plot this to check.

```{r Var_mean_sqaures}
# redefine sample variance calculated from mean square
PM_MB_means$vi <-
   PM_MB_means$Y_Msquare /
   (PM_MB_means$n * PM_MB_means$grain_yield.t.ha ^ 2)

PM_MB_means %>%
   ggplot(aes(x = grain_yield.t.ha, y = vi)) +
   geom_point()
```

@@

Using the model without dose produces a lower AIC, however we should accept the data with dose as there is a significant difference between the models and we should accept the more complicated model `m2`. 
We will retain dose in the trial factor.

```{r}
slimmer_PM_dat <- slimmer_PM_dat[,colnames(slimmer_PM_dat) != "trial_noDose"] #remove second trial column
write.csv(slimmer_PM_dat, file = "cache/slimmer_PM_clusterdat.csv", row.names = FALSE)
```
